---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r echo=FALSE}
library(RMySQL)
db_user <- 'cs5200'
db_password <- '12345678'
db_name <- 'SandboxDB' 
db_host <- 'cs5211-dbs.cbw6wxrvp4bq.us-east-2.rds.amazonaws.com'
db_port <- 3306
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password, 
                   dbname = db_name, host = db_host, port = db_port)

```

```{sql connection=mydb}
drop table if exists incidents;
```


```{sql connection=mydb}
drop table if exists airports;
```

```{sql connection=mydb}
drop table if exists airlines;
```

1. Create Table
```{sql connection=mydb}
create table airlines(
  aid int not null primary key,
  code varchar(30),
  airline varchar(40)
);
```



```{sql connection=mydb}
create table airports (
  pid varchar(80) not null primary key, 
  code  varchar(30), 
  name varchar(80),
  city  varchar(40),
  state  varchar(40),
  country  varchar(40)
);

```

```{sql connection=mydb}
create table incidents(
  iid int  not null  primary key,
  date date,
  depPort varchar(60),
  arrPort varchar(60),
  airline int, 
  aircraft  varchar(60),
  flightPhase  varchar(30)  check(flightPhase in('takeoff', 'landing', 'inflight', 'unknown')), 
  impact  varchar(30) check(impact in('Caused damage','No damage')), 
  foreign key(depPort) references airports(pid),
  foreign key(arrPort) references airports(pid),
  foreign key(airline) references airlines(aid)
);

```


Q2.Read Table and Insert Value
```{r warning=F, message=F}
library(readr)
library(sqldf)
library(lubridate)
BirdStrikesData <- read_csv("BirdStrikesData.csv")
BirdStrikesData$date <- date(mdy_hm(BirdStrikesData$FlightDate))
```

Q2.Read Table and Insert Value
```{r warning=F, message=F}
airlines <- data.frame(
  aid=1:length(unique(BirdStrikesData$`Aircraft: Airline/Operator`)),
  code=NA,
  airline=unique(BirdStrikesData$`Aircraft: Airline/Operator`))

```
Q2.Read Table and Insert Value
```{r}
airlines$code <- as.character(airlines$code)
dbExecute(mydb, sqlAppendTable(mydb, "airlines", airlines, row.names = NULL))
```
Q2.Read Table and Insert Value
```{r warning=F, message=F}
library(tidyverse)
BirdStrikesData <- left_join(BirdStrikesData, airlines, by=c("Aircraft: Airline/Operator"="airline"))
airports <- BirdStrikesData %>% select(`Airport: Name`, `Origin State`) %>%
  mutate(code=NA, city=NA, country=NA)
colnames(airports) <- c("name", "state", "code", "city", "country")
airports <- airports[!duplicated(airports$name), ]
airports$pid <- 1:nrow(airports)
airports <- airports[, c("pid", "code", "name", "city", "state",  "country")]
dbExecute(mydb, sqlAppendTable(mydb, "airports", airports, row.names = NULL))
```

Q2.Read Table and Insert Value
```{r}
BirdStrikesData <- left_join(BirdStrikesData, airports, by=c("Airport: Name"="name"))
```
Q2.Read Table and Insert Value and fix flight phase
```{r}
BirdStrikesData$flightPhase <- rep("unknown", nrow(BirdStrikesData))
BirdStrikesData$flightPhase[BirdStrikesData$`When: Phase of flight`=="Approach"] <- "inflight"
BirdStrikesData$flightPhase[BirdStrikesData$`When: Phase of flight`=="Climb"] <- "takeoff"
BirdStrikesData$flightPhase[BirdStrikesData$`When: Phase of flight`=="Descent"] <- "inflight"
BirdStrikesData$flightPhase[BirdStrikesData$`When: Phase of flight`=="Landing Roll"] <- "landing"
BirdStrikesData$flightPhase[BirdStrikesData$`When: Phase of flight`=="Parked"] <- "landing"
BirdStrikesData$flightPhase[BirdStrikesData$`When: Phase of flight`=="Take-off run"] <- "takeoff"
```

Q2.Read Table and Insert Value Assign synthetic key values to aid, iid, and pid and use them as primary keys.
```{r}
BirdStrikesData$iid <- 1:nrow(BirdStrikesData)
BirdStrikesData$aircraft <- BirdStrikesData$`Aircraft: Make/Model`
BirdStrikesData$impact <- BirdStrikesData$`Effect: Indicated Damage`
BirdStrikesData$depPort <- BirdStrikesData$pid
BirdStrikesData$arrPort <- BirdStrikesData$pid
BirdStrikesData$airline <- BirdStrikesData$aid
incidents <- BirdStrikesData %>% select(iid, date, depPort, arrPort, airline, aircraft,
                                        flightPhase, impact
                                        )
incidents$date <- as.character(incidents$date)
dbExecute(mydb, sqlAppendTable(mydb, "incidents", incidents, row.names = NULL))
```

Q3. Show that the loading of the data
```{sql connection=mydb}
select date,aircraft,flightPhase,impact from incidents

```

Q3.Show that the loading of the data
```{sql connection=mydb}
select pid, name,state from airports
```

Q3.Show that the loading of the data

```{sql connection=mydb}
select aid, airline  from airlines
```

Q4.
```{sql connection=mydb}
select airlines.airline,count(*)
from airlines inner join incidents on airlines.aid=incidents.airline
inner join airports on incidents.arrPort=airports.pid
where airports.name='LaGuardia NY' and flightPhase='landing'
group by airlines.airline;
```

Q5.
```{sql connection=mydb}
select  tt.name
from
(
 select t.name, count(*) as nums
 from 
 (
  select airports.name,incidents.iid
  from airlines inner join incidents on airlines.aid=incidents.airline
  inner join airports on incidents.depPort=airports.pid
  where airlines.airline not in('Business','PRIVATELY OWNED','MILITARY')  
  union all
  select airports.name,incidents.iid
  from airlines inner join incidents on airlines.aid=incidents.airline
  inner join airports on incidents.arrPort=airports.pid
  where airlines.airline not in('Business','PRIVATELY OWNED','MILITARY')  
 )  t
 group by t.name
    order by  count(*)  desc
    limit 1
) tt;
```

Q6.
```{r}
library(sqldf)
library(RSQLite)
fpath = "/Users/diyang/Desktop/CS5200/00/"
dbfile = "BirdStrikesData.csv"
```

```{r}
dbDisconnect(dbcon)
```
Q6. find the number of bird strike incidents by year
```{r}
sqlCmd = "select year(date) as year,count(*) as num
from incidents
group by year(date);"
rs = dbGetQuery(mydb,sqlCmd)
rs
```

Q7. Firsrt find the bird strikes incidents per year from 2005 to 2011
```{r}
sqlCmd = "select year(date) as year,count(*) as sumNumberOfBirdStrike
from incidents where year(date) between '2005' and '2011'
group by year(date);"
df = dbGetQuery(mydb,sqlCmd)
df
```

Q7. creat line plot for bird strikes incidents per year from 2005 to 2011
```{r}
library(ggplot2)
library(lubridate)
p<-ggplot(df,aes(x=year,y=sumNumberOfBirdStrike)) + geom_line()+geom_point() + ggtitle("SumNumberOfBirdStrike")
p
```


Q8. removes a bird strike incident from the database

```{sql connection=mydb}

create procedure proc_1(in airline_name varchar(100))
begin
  delete from incidents
     where airline in (select aid
      from airlines
                        where airline=airline_name);
end ;
call proc_1('AMERICAN AIRLINES'); 

```



